---
title: "Crash Data: City of Tempe"
output: 
  flexdashboard::flex_dashboard:
    theme: flatly
    source: embed
    smart: false
runtime: shiny
---

```{r include = FALSE}

# LOAD PACKAGES

library( DT)
library( ggmap )
library( shiny )
library( knitr )
library( pander )
library( leaflet )
library( viridis )
library( rsconnect )
library( tidyverse )
library( flexdashboard )



# READ IN DATA

url <- paste0( "https://github.com/DS4PS/Data-",
               "Science-Class/blob/master/DATA",
               "/TempeTrafficAccidents.rds?raw=true" )

dat <- readRDS( gzcon( url( url ) ) )



# DATA PREPROCESSING I: INJURIES & FATALITIES

dat        <- na.omit( dat )                                       # Remove NA
dat$fatal  <- dat$Totalfatalities > 0                              # 1+ fatalities
dat$inj    <- dat$Totalinjuries > 0 & dat$Totalfatalities == 0     # 1+ injuries, 0 fatal
dat$nohurt <- dat$Totalfatalities + dat$Totalinjuries == 0         # Harmless

date.vec   <- strptime( dat$DateTime, format = "%m/%d/%y %H:%M" )  # Datetime variables
dat$hour   <- format( date.vec, format = "%H" ) %>% as.numeric()
dat$month  <- format( date.vec, format = "%b" )
dat$day    <- format( date.vec, format = "%a" )
dat$day365 <- format( date.vec, format = "%j" )
dat$week   <- format( date.vec, format = "%V" )



# DATA PREPROCESSING II: NAMED INTERVALS OF TIME

dat <- dat %>% 
  mutate( time.of.day = case_when( hour >= 6 & hour <= 9 ~ "Morning Commute", 
                                   hour >= 16 & hour <= 19 ~ "Evening Commute", 
                                   hour >= 14 & hour <= 15 ~ "School Pickup", 
                                   hour >= 9 & hour <= 13 ~ "Work", 
                                   hour >= 20 & hour <= 23 ~ "Night", 
                                   hour <= 5 & hour >= 0 ~ "Midnight to Dawn" ) )

dat <- dat %>% 
  mutate( day.cat = case_when( day == "Mon" |                                 ## Created a case for mutate by day
                               day == "Tue" |                                 ## All weekdays become 'Weekdays'
                               day == "Wed" |                                 ## All Weekend days become 'Weekend'
                               day == "Thu" |                                 ## This is for use with Drugs & Alcohol Tab
                               day == "Fri" ~ "Weekdays", 
                               day == "Sat" | 
                               day == "Sun" ~ "Weekend" ) )

dat$harm <- ifelse( test = dat$Totalinjuries > 0 | dat$Totalfatalities > 0, 
                    yes = "Harm", 
                    no = "No Harm" )

# DATA PREPROCESSING III: PERMUTATIONS OF INEBRIATION

dat <- dat %>% 
  mutate( d1.substance = case_when( AlcoholUse_Drv1 == "Alcohol" & 
                                    DrugUse_Drv1 == "No Apparent Influence" ~ "Alcohol", 
                                    AlcoholUse_Drv1 == "No Apparent Influence" & 
                                    DrugUse_Drv1 == "Drugs" ~ "Drugs", 
                                    AlcoholUse_Drv1 == "Alcohol" & 
                                    DrugUse_Drv1 == "Drugs" ~ "Alcohol and Drugs", 
                                    AlcoholUse_Drv1 == "No Apparent Influence" & 
                                    DrugUse_Drv1 == "No Apparent Influence" ~ "No Apparent Influence" ) )

dat <- dat %>% 
  mutate( d2.substance = case_when( AlcoholUse_Drv2 == "Alcohol" & 
                                    DrugUse_Drv2 == "No Apparent Influence" ~ "Alcohol", 
                                    AlcoholUse_Drv2 == "No Apparent Influence" & 
                                    DrugUse_Drv2 == "Drugs" ~ "Drugs", 
                                    AlcoholUse_Drv2 == "Alcohol" & 
                                    DrugUse_Drv2 == "Drugs" ~ "Alcohol and Drugs", 
                                    AlcoholUse_Drv2 == "No Apparent Influence" & 
                                    DrugUse_Drv2 == "No Apparent Influence" ~ "No Apparent Influence" ) )

# DATA PREPROCESSING IV: AGE CATEGORIES

dat$age.cat <- case_when( dat$Age_Drv1 >= 0 & 
                          dat$Age_Drv1 <= 18 ~ "Youth", 
                          dat$Age_Drv1 >= 19 & 
                          dat$Age_Drv1 <= 25 ~ "Young Adult", 
                          dat$Age_Drv1 >= 26 & 
                          dat$Age_Drv1 <= 64 ~ "Adult", 
                          dat$Age_Drv1 >= 65 ~ "Senior" )

```

Day & Time
=====================================  

Inputs {.sidebar}
-------------------------------------
Day and Time

```{r}

checkboxGroupInput( "days", 
                    label = h3( "Day of Week" ), 
    choices = list( "Monday"    = "Mon", 
                    "Tuesday"   = "Tue", 
                    "Wednesday" = "Wed", 
                    "Thursday"  = "Thu",
                    "Friday"    = "Fri",
                    "Saturday"  = "Sat",
                    "Sunday"    = "Sun"  ),
    selected = c( "Fri",
                  "Sat",
                  "Sun" ) )

sliderInput( inputId = "hour", 
             label = h3( "Time of Day" ), 
             min = 0, 
             max = 23, 
             value = c( 6, 12 ) )

```

Outputs
-------------------------------------

### Traffic Accidents By Day and Time

```{r}

# LEAFLET MAPPING

renderLeaflet({
  
  d2 <- dat %>%                                                         ## Set dat into a new object to protect changes
    filter( day %in% input$days, 
            hour >= input$hour[1],
            hour <= input$hour[2] )
  
  d2$col.vec <- ifelse( test = d2$nohurt,                               ## Changed every dat$???? instance to d2$????
                        yes = "gray20", 
                        no = ifelse( test = d2$inj, 
                                     yes = "steelblue", 
                                     no = "darkorange" ) ) 
  
  point.size <- d2$Totalinjuries + d2$Totalfatalities                   ## Kept all equations, objects, otherwise the same in tab

  crash.details <- paste0( "Time: ", d2$DateTime, "<br>",
                           "Total Fatalities: ", d2$Totalfatalities, "<br>",
                           "Total Injuries: ", d2$Totalinjuries, "<br>",
                           "Collision type: ", d2$Collisionmanner )
  
  tempe <- leaflet() %>% 
    addProviderTiles( "CartoDB.Positron" )  %>%
    setView( lng = -111.9278, 
             lat = 33.39951, 
             zoom = 13 )
  
  addCircles(tempe, 
             lng = d2$Longitude, 
             lat = d2$Latitude,
             fillColor = d2$col.vec, 
             fillOpacity = 0.5, 
             stroke = FALSE, 
             radius = 50 * ( 1 + 0.33 * point.size ),
             popup = crash.details )

})

```   

Age, Gender, & Method {data-orientation=rows}
=====================================  

Sidebar {.sidebar}
-------------------------------------
Driver Characteristics

```{r}

sliderInput( inputId = "d1age",                                    ## Instructions questions this tab makeup
             label = h4( "Driver 1 Age" ),                         ## but I thought it was Okay considering
             min = 15,                                             ## what I intended to do with my custom tabs
             max = 100,  
             value = c( 18,36 ) )

sliderInput( inputId = "d2age", 
             label = h4( "Driver 2 Age" ), 
             min = 15, 
             max = 100, 
             value = c( 18,36 ) )

selectInput( inputId = "d1gender", 
             label = h4("Driver 1 Gender"), 
             choices = c( "Male",
                          "Female" ), 
             selected = "Female" )

selectInput( inputId = "d2gender", 
             label = h4( "Driver 2 Gender" ), 
             choices = c( "Male",
                          "Female" ), 
             selected = "Female" )

radioButtons( inputId = "d1pedcy", 
              label = h4( "Driver 1 Transportation" ),
              choices = c( "Driver", 
                           "Pedalcyclist", 
                           "Pedestrian" ), 
              selected = "Driver")

radioButtons( inputId = "d2pedcy", 
              label = h4( "Driver 2 Transportation" ),
              choices = c( "Driver", 
                           "Pedalcyclist", 
                           "Pedestrian" ), 
              selected = "Driver" )

```

Row 
-------------------------------------

### Number of Crashes
  
```{r}

renderValueBox({
  
  d2 <- dat %>%                                               ## Another instance of keeping in the d2 object
    filter( Age_Drv1 >= input$d1age[1],
            Age_Drv1 <= input$d1age[2], 
            Age_Drv2 >= input$d2age[1],
            Age_Drv2 <= input$d2age[2], 
            Gender_Drv1 %in% input$d1gender, 
            Gender_Drv2 %in% input$d2gender, 
            Unittype_One %in% input$d1pedcy, 
            Unittype_Two %in% input$d2pedcy )
  
  crashes <- count( d2 )
  
  valueBox( crashes, 
            icon = "fa-car-burst",                             ## Changed icon to fa-car-burst for thematic purpose
            color = ifelse(test = crashes > 50, 
                           yes = "danger", 
                           no = "primary" ) )
  
})

```

### Total Injuries
  
```{r}

renderValueBox({
  
  d2 <- dat %>%                                                        ## Another instance of keeping in the d2 object
    filter( Age_Drv1 >= input$d1age[1],
            Age_Drv1 <= input$d1age[2], 
            Age_Drv2 >= input$d2age[1],
            Age_Drv2 <= input$d2age[2], 
            Gender_Drv1 %in% input$d1gender, 
            Gender_Drv2 %in% input$d2gender, 
            Unittype_One %in% input$d1pedcy, 
            Unittype_Two %in% input$d2pedcy )
  
  total.injuries <- sum( d2$Totalinjuries )
  
  valueBox( total.injuries, 
            icon = "fa-user-injured",                                   ## Changed icon to fa-user-injured, seemed fitting
            color = ifelse(test = total.injuries > 30, 
                           yes = "danger", 
                           no = "primary"  ) )

  })

```

### Total Fatalities
  
```{r}

renderValueBox({
  
  d2 <- dat %>%                                               ## Another instance of keeping in the d2 object
    filter( Age_Drv1 >= input$d1age[1],
            Age_Drv1 <= input$d1age[2], 
            Age_Drv2 >= input$d2age[1],
            Age_Drv2 <= input$d2age[2], 
            Gender_Drv1 %in% input$d1gender, 
            Gender_Drv2 %in% input$d2gender, 
            Unittype_One %in% input$d1pedcy, 
            Unittype_Two %in% input$d2pedcy )

  total.fatalities <- sum( d2$Totalfatalities )
  
  valueBox( total.fatalities, 
            icon = "fa-skull-crossbones",                     ## Changed icon to fa-skull-crossbones for fatality
            color = ifelse(test = total.fatalities > 5, 
                           yes = "danger", 
                           no = "primary" ) )
  
})

```

### Rate of Harm
  
```{r}

renderValueBox({
  
  d2 <- dat %>%                                                 ## Another instance of keeping in the d2 object
    filter( Age_Drv1 >= input$d1age[1],
            Age_Drv1 <= input$d1age[2], 
            Age_Drv2 >= input$d2age[1],
            Age_Drv2 <= input$d2age[2], 
            Gender_Drv1 %in% input$d1gender, 
            Gender_Drv2 %in% input$d2gender, 
            Unittype_One %in% input$d1pedcy, 
            Unittype_Two %in% input$d2pedcy )
  
  rate.of.harm <- round( length( which( d2$harm == "Harm" ) ) / count( d2 ), 3 )    ## Liked this rate, kept it throughout
  
  valueBox( rate.of.harm,                                                  
            icon = "fa-calculator",                                          ## Changed icon to fa-calculator for mathiness
            color = ifelse( test = rate.of.harm > 0.45, 
                            yes = "danger", 
                            no = "primary" ) )
  
})

```

Outputs
-------------------------------------

### Traffic Accidents by Driver Characteristics

```{r}

renderLeaflet({
  
  d2 <- dat %>%                                             ## Another instance of keeping in the d2 object
    filter( Age_Drv1 >= input$d1age[1],
            Age_Drv1 <= input$d1age[2], 
            Age_Drv2 >= input$d2age[1],
            Age_Drv2 <= input$d2age[2], 
            Gender_Drv1 %in% input$d1gender, 
            Gender_Drv2 %in% input$d2gender, 
            Unittype_One %in% input$d1pedcy, 
            Unittype_Two %in% input$d2pedcy )
  
  d2$col.vec <- ifelse( test = d2$nohurt, 
                        yes = "gray20", 
                        no = ifelse(test = d2$inj, 
                                   yes = "steelblue", 
                                   no = "darkorange" ) )              
    
  point.size <- d2$Totalinjuries + d2$Totalfatalities

  crash.details <- paste0( "Time: ", d2$DateTime, "<br>",
                           "Total Fatalities: ", d2$Totalfatalities, "<br>",
                           "Total Injuries: ", d2$Totalinjuries, "<br>",
                           "Collision type: ", d2$Collisionmanner )
  
  tempe <- leaflet() %>% 
    addProviderTiles( "CartoDB.Positron" )  %>%
    setView( lng = -111.9278, 
             lat = 33.39951, 
             zoom = 13 )
  
  
  addCircles( tempe, 
             lng = d2$Longitude, 
             lat = d2$Latitude,
             fillColor = d2$col.vec, 
             fillOpacity = 0.5, 
             stroke = FALSE, 
             radius = 50 * ( 1 + 0.33 * point.size ),
             popup = crash.details )


})

```   

Comparisons {data-orientation=rows}
=====================================  

Sidebar {.sidebar}
-------------------------------------
Driver Characteristics

```{r}

sliderInput( inputId = "d1age1",                                ## This tab was easily mixed up with tab 1
             label = h4( "Driver 1A Age" ),                     ## So needed to be isolated to data within tab 2
             min = 15,                                          ## Created new inputId's for each sliderInput
             max = 100,                                         ## and set the data points for tab opening
             value = c( 15, 21 ) )

sliderInput( inputId = "d1age2",                                ## I went with a comparison within Driver 1 group
             label = h4( "Driver 1B Age" ),                     ## so I made separate inputs for Driver 1A and 1B
             min = 15,                                          ## I liked how the widgets were measuring and kept
             max = 100,                                         ## the nuts and bolts of it, but modified titles
             value = c( 36, 54 ) )

selectInput( inputId = "d1gender1", 
             label = h4( "Driver 1A Gender" ), 
             choices = c( "Male",
                         "Female" ), 
             selected = "Male" )

selectInput( inputId = "d1gender2", 
             label = h4( "Driver 1B Gender" ), 
             choices = c( "Male",
                         "Female" ), 
             selected = "Male" )

radioButtons( inputId = "d1pedcy1",                              ## This is the only tab with vehicle type
              label = h4( "Driver 1A Transportation" ),
              choices = c( "Driver",
                           "Pedalcyclist",
                           "Pedestrian" ), 
              selected = "Driver" )

radioButtons( inputId = "d1pedcy2", 
              label = h4( "Driver 1B Transportation" ),
              choices = c( "Driver",
                          "Pedalcyclist",
                          "Pedestrian" ), 
              selected = "Driver" )

```

Row 
-------------------------------------

### Total Crashes
  
```{r}

renderValueBox({
  
  d11 <- dat %>%                                                       ## Isolated dat object for Driver 1A
    filter( Age_Drv1 >= input$d1age1[1],                               ## had to correctly link inputs in widgets
            Age_Drv1 <= input$d1age1[2], 
            Gender_Drv1 %in% input$d1gender1, 
            Unittype_One %in% input$d1pedcy1 )
  
  crashes <- count( d11 )                                              ## Kept to object d11 for Driver 1A
  
  valueBox( crashes, 
            icon = "fa-a",                                             ## Specialized A icon for Driver A crashes
            color = ifelse( crashes > 50, "danger", "primary" ) )
  
})

```

### Total Injuries
  
```{r}

renderValueBox({
  
  d11 <- dat %>%                                                         ## Kept to object d11 for Driver 1A
    filter( Age_Drv1 >= input$d1age1[1],
            Age_Drv1 <= input$d1age1[2], 
            Gender_Drv1 %in% input$d1gender1, 
            Unittype_One %in% input$d1pedcy1 )
  
  total.injuries <- sum( d11$Totalinjuries )
  
  valueBox( total.injuries, 
            icon = "fa-user-injured",                                     ## Changed icon to fa-user-injured for theme
            color = ifelse(total.injuries > 30, "danger", "primary" ) )
  
})

```

### Total Crashes
  
```{r}

renderValueBox({
  
  d12 <- dat %>%                                                      ## Isolated dat object for Driver 1B
    filter( Age_Drv1 >= input$d1age2[1],
            Age_Drv1 <= input$d1age2[2],                              ## Had to correctly link inputs in widgets
            Gender_Drv1 %in% input$d1gender2, 
            Unittype_One %in% input$d1pedcy2 )
  
  crashes <- count( d12 )
  
  valueBox( crashes, 
            icon = "fa-b",                                             ## Specialized B letter icon for Driver 1 B
            color = ifelse( crashes > 50, "danger", "primary" ) )
  
})

```

### Total Injuries
  
```{r}

renderValueBox({
  
  d12 <- dat %>%                                                        ## Kept to object d12 for Driver 1B
    filter( Age_Drv1 >= input$d1age2[1],
            Age_Drv1 <= input$d1age2[2], 
            Gender_Drv1 %in% input$d1gender2, 
            Unittype_One %in% input$d1pedcy2 )
  
  total.injuries <- sum( d12$Totalinjuries )
  
  valueBox( total.injuries, 
            icon = "fa-user-injured",                                    ## icon becomes fa-user-injured for injuries
            color = ifelse(total.injuries > 30, "danger", "primary" ) )
  
})

```

Column
-------------------------------------

### Traffic Accidents for Driver 1A 

```{r}

renderLeaflet({                                                 ## Columns begin for Driver 1 Maps
  
  d11 <- dat %>%                                                ## Kept to object d11 for Driver 1A
    filter( Age_Drv1 >= input$d1age1[1],
            Age_Drv1 <= input$d1age1[2], 
            Gender_Drv1 %in% input$d1gender1, 
            Unittype_One %in% input$d1pedcy1 )
  
  d11$col.vec <- ifelse( test = d11$nohurt,                      ## Had to change several objects to match d11
                         yes = "gray20", 
                         no = ifelse( test = d11$inj, 
                                      yes = "steelblue", 
                                      no = "darkorange" ) )              
    
  point.size <- d11$Totalinjuries + d11$Totalfatalities

  crash.details <- paste0( "Time: ", d11$DateTime, "<br>",
                           "Total Fatalities: ", d11$Totalfatalities, "<br>",
                           "Total Injuries: ", d11$Totalinjuries, "<br>",
                           "Collision type: ", d11$Collisionmanner )
  
  tempe <- leaflet() %>% 
    addProviderTiles( "CartoDB.Positron" ) %>%
    setView( lng = -111.9278, 
             lat = 33.39951, 
             zoom = 13 )
  
  addCircles( tempe, 
              lng = d11$Longitude, 
              lat = d11$Latitude,
              fillColor = d11$col.vec, 
              fillOpacity = 0.5, 
              stroke = FALSE, 
              radius = 50 * ( 1 + 0.33 * point.size ),
              popup = crash.details )


})
```   

### Traffic Accidents for Driver 1B

```{r}

renderLeaflet({                                                 ## Another column for Driver 1 Maps
  
  d12 <- dat %>%                                                ## Kept to object d12 for Driver 1B
    filter( Age_Drv1 >= input$d1age2[1],
            Age_Drv1 <= input$d1age2[2], 
            Gender_Drv1 %in% input$d1gender2, 
            Unittype_One %in% input$d1pedcy2 )
  
  d12$col.vec <- ifelse( test = d12$nohurt,                     ## Had to change several objects to match d12
                         yes = "gray20", 
                         no = ifelse(test = d12$inj, 
                                    yes = "steelblue", 
                                    no = "darkorange" ))
  
  point.size2 <- d12$Totalinjuries + d12$Totalfatalities

  crash.details2 <- paste0( "Time: ", d12$DateTime, "<br>",    ## Another instance of keeping in isolated tab dat
                            "Total Fatalities: ", d12$Totalfatalities, "<br>",
                            "Total Injuries: ", d12$Totalinjuries, "<br>",
                            "Collision type: ", d12$Collisionmanner )
  
  tempe2 <- leaflet() %>% 
    addProviderTiles( "CartoDB.Positron" ) %>%
    setView( lng = -111.9278, 
             lat = 33.39951, 
             zoom = 13 )
  
  addCircles( tempe2, 
              lng = d12$Longitude, 
              lat = d12$Latitude,
              fillColor = d12$col.vec, 
              fillOpacity = 0.5, 
              stroke = FALSE, 
              radius = 50 * ( 1 + 0.33 * point.size2 ),
              popup = crash.details2 )

})

```   


Collision Type {data-orientation=rows}
=====================================  

Sidebar {.sidebar}
-------------------------------------
Accident Characteristics

```{r}

radioButtons( inputId = "collision.type",                                 # Beginning of custom tab for Collision Type
              label = h4("Collision Type"),                               # Established labels and inputs
              choices = sort(c("Rear To Rear",                            # Sorted list of data Collisionmanner names
                         "Unknown",
                         "Sideswipe Opposite Direction",
                         "Rear To Side",
                         "Head On",
                         "Other",
                         "Sideswipe Same Direction",
                         "ANGLE (Front To Side)(Other Than Left Turn)",
                         "Left Turn",
                         "Rear End" ) ), 
              selected = "Head On" )                                      # Selected Head On for tab open, reminds me
                                                                          # of a Pixies song on Trompe Le Monde
checkboxGroupInput( "timeofday", 
                   label = h3( "Time of Day" ),                           # Wanted to use a mutated hour span checkbox
    choices = list( "Morning Commute",                                    # I used the casewhen given but unused in the 
                    "Evening Commute",                                    # dashboard foundation that was provided 
                    "School Pickup", 
                    "Work",
                    "Night",
                    "Midnight to Dawn" ),
    selected = c( "Morning Commute",                    
                  "Evening Commute" ) )
                                                                          # Provided a disclaimer for ERROR possibilities
h6( em( "Disclaimer: At least one 'Time of Day' box must be checked to function properly." ), style="color:green" ) 

```

Row 
-------------------------------------

### Total Crash Type

```{r}

renderValueBox({
  
  d20 <- dat %>%                                                    # Set dat into a new object isolated for this tab
         filter( time.of.day %in% input$timeofday,                  # Fixed inputs correctly for widgets
                 Collisionmanner %in% input$collision.type )
  
  type.crashes <- count( d20 )                                      # Another instance of using one object per tab dat
  
  valueBox( type.crashes, 
            icon = "fa-car-burst",                                             # Changed icon to fa-car-burst because cars
            color = ifelse( type.crashes > 150, "danger", "primary" ) )        # Modified number for danger scale
  
})

```


### Total Injuries
  
```{r}

renderValueBox({
  
  d20 <- dat %>%                                                # Another instance of using exclusive object per tab dat
         filter( time.of.day %in% input$timeofday, 
                 Collisionmanner %in% input$collision.type )
  
  total.injuries <- sum( d20$Totalinjuries )
  
  valueBox( total.injuries, 
            icon = "fa-user-injured",                           # Changed icon to fa-user-injured
            color = ifelse( test = total.injuries > 30,         # Modified number for danger scale
                            yes = "danger", 
                            no = "primary"  ) )
 
  })

```

### Total Fatalities
  
```{r}

renderValueBox({
  
  d20 <- dat %>%                                                 # Another instance of using exclusive object per tab dat
         filter( time.of.day %in% input$timeofday,
                 Collisionmanner %in% input$collision.type )

  total.fatalities <- sum( d20$Totalfatalities )
  
  valueBox( total.fatalities, 
            icon = "fa-skull-crossbones",                        # fa-skull-crossbones used as icon for fatalities
            color = ifelse( test = total.fatalities > 5,         # Total fatalities scale to better match widgets
                            yes = "danger", 
                            no = "primary" ) )
  
})

```

### Rate of Harm
  
```{r}

renderValueBox({
  
  d20 <- dat %>%                                                 # Another instance of using exclusive object per tab dat
         filter( time.of.day %in% input$timeofday, 
                 Collisionmanner %in% input$collision.type )
  
  rate.of.harm <- round( length( which( d20$harm == "Harm" ) ) / count( d20 ), 3 )
  
  valueBox( rate.of.harm, 
            icon = "fa-calculator",
            color = ifelse( test = rate.of.harm > 0.45, 
                            yes = "danger", 
                            no = "primary" ) )
  
})

```

Outputs
-------------------------------------

### Traffic Accidents by Accident Characteristics

```{r}

renderLeaflet({
  
  d20 <- dat %>%                                                  ## An instance of using exclusive object per tab dat
         filter( time.of.day %in% input$timeofday, 
                 Collisionmanner %in% input$collision.type )
  
  d20$col.vec <- ifelse( test = d20$nohurt, 
                         yes = "gray20", 
                         no = ifelse(test = d20$inj, 
                                     yes = "steelblue", 
                                     no = "darkorange" ) )              
    
  point.size <- d20$Totalinjuries + d20$Totalfatalities

  crash.details <- paste0( "Time: ", d20$DateTime, "<br>",         ## Another instance of keeping in isolated tab dat
                           "Total Fatalities: ", d20$Totalfatalities, "<br>",
                           "Total Injuries: ", d20$Totalinjuries, "<br>",
                           "Collision type: ", d20$Collisionmanner )
  
  tempe <- leaflet() %>% 
    addProviderTiles( "CartoDB.Positron" )  %>%
    setView( lng = -111.9278, 
             lat = 33.39951, 
             zoom = 13 )
  
  
  addCircles( tempe, 
              lng = d20$Longitude, 
              lat = d20$Latitude,
              fillColor = d20$col.vec, 
              fillOpacity = 0.5, 
              stroke = FALSE, 
              radius = 50 * ( 1 + 0.33 * point.size ),
              popup = crash.details )

})

```


Drugs & Alcohol {data-orientation=rows}
=====================================  

Sidebar {.sidebar}
-------------------------------------
Accident Characteristics

```{r}

radioButtons( inputId = "d1.impairment",                        ## Started new tab for Drugs & Alcohol accidents
              label = h4( "Driver 1 Impairment" ),              ## Established input IDs and labels for widgets
              choices = c( "Alcohol",
                           "Drugs",
                           "Alcohol and Drugs",
                           "No Apparent Influence" ), 
              selected = "Alcohol" )                            ## Pulled names from mutated casewhen given 

radioButtons( inputId = "d2.impairment",                        ## Making a driver 1 and driver 2 radio option
              label = h4( "Driver 2 Impairment" ),
              choices = c( "Alcohol",
                           "Drugs",
                           "Alcohol and Drugs",
                           "No Apparent Influence" ), 
              selected = "No Apparent Influence" )              ## Tab open shows D1 Alcohol and D2 No Apparent Influence

sliderInput( inputId = "hour1",                                 ## Time of Day slider input fits well in the widget space
             label = h3( "Time of Day" ), 
             min = 0, 
             max = 23, 
             value = c( 0, 23 ) )                              

checkboxGroupInput( "days1",                                    ## Created my own casewhen for weekends / weekdays
                    label = h3( "Part of Week" ),               ## Called Part of Week
                    choices = list( "Weekdays", 
                                    "Weekend" ),
                    selected = "Weekend" )                      ## Set tab open to Weekend drivers

h6( em( "Disclaimer: Combinations of choices with insufficient data result in ERROR." ), style="color:red" )
                                                                ## Created disclaimer to notify users of potential error
```

Row 
-------------------------------------

### Inebriation Crash Total

```{r}

renderValueBox({
  
  d30 <- dat %>%                                                 ## Established a new object for this filtered dat tab
    filter( d1.substance %in% input$d1.impairment,               ## Pulled from substance use in accident reports 
            d2.substance %in% input$d2.impairment,               ## Substance is a mutated casewhen found in base code
            hour >= input$hour1[1],                              
            hour <= input$hour1[2],    
            day.cat %in% input$days1 )                           ## Made use of my self-made Part of Week categories 
  
  type.crashes <- count( d30 )
  
  valueBox( type.crashes, 
            icon = "fa-beer-mug-empty",                                   ## Went with the beer mug for drunk driving
            color = ifelse( type.crashes > 20, "danger", "primary" ) )
  
})

```


### Total Injuries
  
```{r}

renderValueBox({
  
  d30 <- dat %>%                                               ## Continued using this isolated d30 filtered dat tab
    filter( d1.substance %in% input$d1.impairment,
            d2.substance %in% input$d2.impairment, 
            hour >= input$hour1[1],
            hour <= input$hour1[2],
            day.cat %in% input$days1 )
  
  total.injuries <- sum( d30$Totalinjuries )                   ## Another instance of keeping in isolated tab dat
  
  valueBox( total.injuries, 
            icon = "fa-user-injured",                          ## Changed icon to fa-user-injured for total injuries
            color = ifelse( test = total.injuries > 30, 
                            yes = "danger", 
                            no = "primary" ) )

  })

```

### Total Fatalities
  
```{r}

renderValueBox({
  
  d30 <- dat %>%                                                  ## Another instance of keeping in isolated tab dat
    filter( d1.substance %in% input$d1.impairment,
            d2.substance %in% input$d2.impairment,
            hour >= input$hour1[1],
            hour <= input$hour1[2],
            day.cat %in% input$days1 )

  total.fatalities <- sum( d30$Totalfatalities )                  ## Another instance of keeping in isolated tab dat
  
  valueBox( total.fatalities, 
            icon = "fa-skull-crossbones",                         ## fa-skull-crossbones icon for fatalities
            color = ifelse( test = total.fatalities > 5,           
                            yes = "danger", 
                            no = "primary" ) )
  
})

```

### Rate of Harm
  
```{r}

renderValueBox({
  
  d30 <- dat %>%                                         ## Another instance of keeping in isolated tab dat
    filter( d1.substance %in% input$d1.impairment,
            d2.substance %in% input$d2.impairment,
            hour >= input$hour1[1],
            hour <= input$hour1[2],
            day.cat %in% input$days1 )
  
  rate.of.harm <- round( length( which( d30$harm == "Harm" ) ) / count( d30 ), 3 )
                                                          ## Another instance of keeping in isolated tab dat
  valueBox( rate.of.harm, 
            icon = "fa-calculator",                       ## fa-calculator icon used for mathiness
            color = ifelse(test = rate.of.harm > 0.45,    ## lowered rate of harm number for fitness
                           yes = "danger", 
                           no = "primary") )
  
})

```

Row 
-------------------------------------

### Traffic Accidents by Accident Characteristics

```{r}

renderLeaflet({
  
  d30 <- dat %>%                                                   ## Another instance of keeping in isolated tab dat
    filter( d1.substance %in% input$d1.impairment,
            d2.substance %in% input$d2.impairment,
            hour >= input$hour1[1],
            hour <= input$hour1[2],
            day.cat %in% input$days1 )
   
  d30$col.vec <- ifelse( test = d30$nohurt,                        ## Another instance of keeping in isolated tab dat
                         yes = "gray20", 
                         no = ifelse(test = d30$inj,               ## Another instance of keeping in isolated tab dat
                                     yes = "steelblue", 
                                     no = "darkorange" ) )              
    
  point.size <- d30$Totalinjuries + d30$Totalfatalities

  crash.details <- paste0( "Time: ", d30$DateTime, "<br>",
                           "Total Fatalities: ", d30$Totalfatalities, "<br>",
                           "Total Injuries: ", d30$Totalinjuries, "<br>",
                           "Collision type: ", d30$Collisionmanner )
  
  tempe <- leaflet() %>% 
    addProviderTiles( "CartoDB.Positron" )  %>%
    setView( lng = -111.9278, 
             lat = 33.39951, 
             zoom = 13 )
  
  
  addCircles( tempe, 
              lng = d30$Longitude,                                ## Another instance of keeping in isolated tab dat
              lat = d30$Latitude,
              fillColor = d30$col.vec, 
              fillOpacity = 0.5, 
              stroke = FALSE, 
              radius = 50 * ( 1 + 0.33 * point.size ),
              popup = crash.details )

})

```





About
===================================== 

Row 
-------------------------------------

### About this Dashboard

Source data is Tempe, Arizona Vehicle Accident Data for Years 2012 - 2018. The purpose of this dashboard is to inform and educate regarding the frequency of accidents and driver detail:

* 'DAY & TIME' tab shows users general accident areas on a Tempe map by weekday and time span.
* 'AGE, GENDER, & METHOD' tab user analyzes driver characteristics from the Tempe data and map.
* 'COMPARISONS' allows user to see map and accident details for two same-data, at-fault drivers.
* 'COLLISION TYPE' grants the user a search by manner of collision and shows numbers for harmfulness.
* 'DRUGS & ALCOHOL' provides user with accident characteristics by impairment level of drivers by time span.
* 'ABOUT' offers users various details regarding the data set and dashboard intent, as well as an author bio.
* 'DATA' very straight-forwardly gives the user access to the source table data that this dashboard is built from.
* 'SOURCE CODE' is for anyone interested in reading the Author's raw code that powers this R Shiny dashboard.

When engaging with this dashboard, pay close attention to the harm rate of drug and/or alcohol impairment versus no influence. Also, notice how time of day, day of week, and part of week correlated with harm rate and crash total. Taking special consideration of collision type might help a driver anticipate especially dangerous road and traffic situations. Thank you for your time and attention.

### Dashboard Author

Tyler Naes:

* Originally from Missouri and currently lives with wife, Rose, in Tennessee.
* An ASU Program Evaluation and Data Analytics graduate student with a new-found passion for programming.
* For professional inquiries, contact via email at tbnaes@gmail.com.

Attribution to Dr. Jesse Lecy for the base dashboard

Row
-------------------------------------

DATA DICTIONARY

```{r}

url.dd <- paste0( "https://raw.githubusercontent.com",
                  "/DS4PS/cpp-526-fall-2019/master/l",
                  "abs/final-project/TempeTrafficAcc",
                  "identsDataDictionary.csv" )

data.dictionary <- read.csv( url.dd, 
                             stringsAsFactors = FALSE )

data.dictionary$description <- stringi::stri_trans_general( data.dictionary$description,
                                                            "latin-ascii" )

data.dictionary %>%
  select( column, description ) %>%
  pander()

```

Data
=====================================  

```{r}

these.buttons <- c( "copy", 
                    "csv", 
                    "pdf", 
                    "print" )

renderDataTable({
  
  datatable( dat,
             rownames = FALSE, 
             fillContainer = TRUE, 
             style = "bootstrap",
             class = "table-condensed table-striped",
             extensions = "Buttons", 
             options = list( dom = "Bfrtip", 
                             buttons = these.buttons ) )

  }, server = FALSE )  ## Changed for a full download of CSV, etc.

```
